Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence and Forensic Workflow

### 2009-03-13 paper/paper.tex
More work - added a section describing the resolver.

%\documentclass[10pt,twocolumn]{article}
\documentclass[10pt, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%\documentclass{article}
%\documentclass[12pt]{IEEEtran}
%\documentclass{elsart5p}
\usepackage{listings}
\lstset{language=Python,
	frame=single,
	captionpos=b,
	frameround=tttt,
	basicstyle=\small,          % print whole listing small
	keywordstyle=\color{black}\bfseries,
                            % underlined bold black keywords
	identifierstyle=,           % nothing happens
	commentstyle=\color{white}, % white comments
	stringstyle=\ttfamily,      % typewriter type for strings
	showstringspaces=false}     % no special string spaces

@@

\subsection{Global Distributed Evidence Management}
While Garfinkel designed AFF for use on a single machine that could
both image evidence and perform analysis, the main use of AFF in
practice has been in distributed environments in which imaging and
analysis takes place in multiple locations and is performed by
multiple individuals.

@@

\subsection{New Concepts and Terminology}

@@

\subsection{Streams}
As discussed above, individual segments in an AFF4 evidence set can be
combined into a single logical object called a \emph{stream}. AFF4
supports an arbitrary number of streams per volume. The AFFLIB library
allows the segments in a stream to be operated on as if they were a
single file inside a file system by supporting the traditional
POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
\texttt{read()}.

AFFLIB also supports the \emph{writing} of streams, a feature that is
used by Garfinkel's \texttt{aimage} disk imaging program. In general
arbitrary writing to streams can be expensive, especially when writing
is combined with seek operations. The current \texttt{aimage} program
will create AFF files with gaps in them during its error recovery
process (when \texttt{aimage} seeks to the end of a disk and reads it
backwards in an attempt to get around media failures). We are working
on a rewrite of \texttt{aimage} which performs imaging and recovery
on bevy boundaries, rather than using the current arbitrary
``high-water-mark'' and ``low-water-mark.''

All streams within an Evidence Set \emph{must} have a properties segment. The
properties segment describes what the stream contains and determines
how the stream is processed.

\subsection{ZIP64}
As described above, AFF4 volumes are only required to store and
retrieve segments by their URN. For AFF4, we have changed the default
volume container file format to ZIP64\cite{zipspecs}. There are many
reasons for this decision:

@@

\item ZIP64 libraries are readily available making proprietary implementations of
interfaces to the AFF4 volume format simple to write. For example a
simple python program to dump out an Image stream is illustrated in
Listing \ref{python_code}.

@@

\begin{lstlisting}[
	float=tb, label=python_code, caption=Sample Python code to
	dump out an Image Stream. As can be seen the chunk index
	segment is used to slice the data segment into chunks. The
	chunks are decompressed and written to the output file.]
volume=zipfile.ZipFile(FILE)
outfd = open("dump.dd","w")
bevy_count = 0

while 1:
    idx_segment = volume.read(STREAM + 
		"%08d.idx" % bevy_count)
    segment = volume.read(STREAM + 
		"%08d" % bevy_count)
    indexes = struct.unpack("<" + "l" * \
	(len(idx_segment)/4),idx_segment)

    for i in range(len(indexes)-1):
	chunk = segment[indexes[i]:
	                indexes[i+1]]
        outfd.write(zlib.decompress(chunk))

    bevy_count += 1
\end{lstlisting}

@@

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
  \caption{The basic structure of a Zip archive}
  \label{zip_structure}
  \end{center}
\end{figure}

It is important to note that AFF4 only requires that the underlying storage mechanism
be capable of storing multiple named segments of data. Although our
AFF4 implementation will use the ZIP64 file format as an underlying storage
mechanism, our system also supports legacy AFF1 volumes as well as
segments that are stored in other storage systems---for example,
Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.

\subsubsection{Unused ZIP64 Features}

There are a number of aspects in the Zip file format specifications
which we do not utilize. 
\begin{itemize}
\item We ignore Zip64's built-in support for splitting
archives into multiple ZIP files. Instead, our implementation treats
each volume as complete and stand-alone ZIP file. The AFF4
implementation then considers the segments contained within as 
belonging to a larger collection. The reason behind this decision is that
often it is needed to add new volumes to an existing set without
modifying the existing set at all. Zip64's built-in support for
splitting an archive renders each individual volume in the set
unusable unless all are available.

@@

\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}

In many forensic applications, individual segments within the volume
need to be updated. To this end, Garfinkel implemented several
``update'' functions in AFF1\cite{garfinkel:aff}.

There were two problems with the original AFF updating approach:
\begin{itemize}
\item Updates can cause the volume to
become fragmented as holes appear within the middle of the volume, and
updated segments are added at the end.
\item Updates could be performed silently, raising questions of
  forensic soundness.
\end{itemize}

In AFF4, when a new segment is added to an existing volume, the new
segment header overwrites the volume central directory, and the
central directory should be rewritten including the new entry
immediately after the last file in the archive. If it is not possible
to overwrite the file, the new segment can be written at the end,
followed by an updated copy of the central directory (in that case the
old central directory remains in the file). This is illustrated in
Figure \ref{zip_structure}. Implementations may delay
writing the central directory until the file is closed; should the
system crash before the central directory is written, it can be
recovered by scanning the entire ZIP file.

A segment within the archive is considered to be updated when a new
segment is added to the archive with the same name and a later
timestamp. Implementations satisfy read requests for duplicate
segment names with those segments with the later timestamp, or if the
two timestamps are the same, the segment which appears at a later
offset in the volume. This effectively allows for a segment to be
updated without leading to internal volume fragmentation.

The fact that the older version of all updated segments remains intact
within each volume can be used for auditing, or to allow a rollback to
previous versions of the segment. We have created a new tool,
\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
creating new volumes and redistributing segments among them.

\section{Standard AFF4 Streams}
In this section we discuss specific stream types that we have created
in our initial AFF4 implementation.

\subsection{The Image Stream}
\label{image_stream}
The AFF4 \emph{Image} stream contains a single read-only forensic data
set. For example, this stream might contains a hard
disk image, a memory image or a network capture (in PCAP
format). Random read access into the stream is provided with
AFFLIB. (AFFLIB also provided the ability to randomly \emph{write}
into the evidence stream for sanitization purposes, although this
feature may be removed at a later date.)

@@

The image itself is stored in \emph{pages} within the AFF4 file. Pages are
segments named as an 8 digit, zero padded decimal integer
representation of their chunk id (e.g. ``default/00000032''). Image
streams must specify the \texttt{page\_size} attribute, as the number of
image bytes each chunk contains. Pages may be compressed at the
page level, or may be broken up into multiple chunks (typically 8KB),
and individually compressed. In this case the beginning of the page
segment will contain an offset table indicating the location of each
compression chunk. This is similar to the technique used by the Expert
Witness file format used by EnCase\cite{encase-3.0} and implemented by
the open source libewf\cite{libewf} package.

>>>>

%\documentclass[10pt,twocolumn]{article}
\documentclass[10pt, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%\documentclass{article}
%\documentclass[12pt]{IEEEtran}
%\documentclass{elsart5p}
\usepackage{listings}
\lstset{language=Python,
	frame=single,
	captionpos=b,
	frameround=tttt,
	basicstyle=\scriptsize,
	keywordstyle=\color{black}\bfseries,
                            % underlined bold black keywords
	identifierstyle=,           % nothing happens
	commentstyle=\color{white}, % white comments
	stringstyle=\ttfamily,      % typewriter type for strings
	showstringspaces=false}     % no special string spaces

@@

\section{Global Distributed Evidence Management}
While Garfinkel designed AFF for use on a single machine that could
both image evidence and perform analysis, the main use of AFF in
practice has been in distributed environments in which imaging and
analysis takes place in multiple locations and is performed by
multiple individuals.

@@

@@

\section{Metadata and the Universal Resolver}
The AFF4 model treats metadata as an abstract concept which may exists
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4 universe.

The AFF4 design extends beyond the management of a single volume,
stream or image to a universal system for managing data of many
types. This necessarily means that a single running instance is
generally unable to have visibility of the entire AFF4 universe. For
example, if a volume is opened which contains a Map Stream targeting a
stream stored in a different volume, its not generally possible to
tell in which volume the stream exists, and where that volume is
actually stored.

To provide this global visibility of metadata we define a central
metadata management entity, named the {\em Universal Resolver}. The
Universal Resolver contains all the metadata about the AFF4 universe,
that is to say it is able to resolve queries about any attribute about
any URN in the universe.

Although the resolver has complete visibility of all attributes, it is
still useful to store metadata within the volume itself, particularly
data pertaining to the volume itself. If we did not store the metadata
within the volume itself, then the volume would not be accessible to
implementations which do not have this metadata.

To this end we define a way for serializing metadata statements (or
triples) into a standard format which implementations can load into
their respective resolvers when parsing the volume. Triples can be
stored in segments having a URN ending with ``{\em properties}''. The
AFFLIB implementation loads these segments automatically into the
Universal Resolver.

Triples are stored within the properties segment one per line, with
the subject URN (encoded according to RFC1737), followed by whitespace
and the attribute name. This is then followed by the equal sign and
the UTF8 encoding of the value. An example properties file for an
Image Stream is shown in Listing \ref{image_metadata}.

It is important to stress that the properties file is simply a
serialization of statements into volume segments. The statements may
exist without being stored in a volume (for example, being stored on
an external SQL server). Alternatively, these statements may be stored
in some other way inside or outside the volume (e.g. SQLite
databases).

When the volume is loaded, the AFFLIB implementation automatically
loads any properties files and populates its Universal Resolver with
the information it has visibility of. AFFLIB provides a mechanism to
use an external resolver stored on a MySQL database to provide a
Universal Resolver that shares information between different instances
on the same network.

\begin{lstlisting}[
	caption=Example properties files for several AFF4 objects (URNs are
shortened for illustration).,
	label=image_metadata,
	float=tb,
	 ]
Directory Volume:
  urn:aff2:f901be8e-d4b2 aff2:stored=http://../case1/
  urn:aff2:f901be8e-d4b2 aff2:type=directory

ZipFile Volume:
  urn:aff2:98a6dad6-4918 aff2:stored=file:///file.zip
  urn:aff2:98a6dad6-4918 aff2:type=zip

Image Stream:
  urn:aff2:83a3d6db-85d5 aff2:stored=urn:aff2:f901be8e-d4b2
  urn:aff2:83a3d6db-85d5 aff2:chunks_in_segment=256
  urn:aff2:83a3d6db-85d5 aff2:chunk_size=32k
  urn:aff2:83a3d6db-85d5 aff2:type=image
  urn:aff2:83a3d6db-85d5 aff2:size=5242880

Map Stream:
  urn:aff2:ed8f1e7a-94aa aff2:target_period=3
  urn:aff2:ed8f1e7a-94aa aff2:image_period=6
  urn:aff2:ed8f1e7a-94aa aff2:blocksize=64k
  urn:aff2:ed8f1e7a-94aa aff2:stored=urn:aff2:83a3d6db-85d5
  urn:aff2:ed8f1e7a-94aa aff2:type=map
  urn:aff2:ed8f1e7a-94aa aff2:size=0xA00000

Link Object:
  map aff2:target=urn:aff2:ed8f1e7a-94aa
  map aff2:type=link
\end{lstlisting}

\section{Volumes}
The volume object is responsible for providing storage for
segments. Segments are stored and retrieved using their URNs. We
describe two different implementations of volume objects, namely the
{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
convert from one implementation to another easily, without effecting
any external references. This is because external references are to
the Volume URN itself, so its safe to convert from one representation
to another. In this way its possible for example, to unzip a Zip64
volume into a directory and vice versa.

\subsection{Directory Volumes}
The Directory Volume is the simplest type of volume. It simply stores
different segments based on their URNs in a single directory. Since
some filesystems are unable to represent URNs accurately (e.g. Windows
has many limitations on the types of characters allowed for a
filename), the Directory Volume encodes URNs according to RFC1738
\cite{RFC1738}; non-printable characters are escaped with a \%
followed by the ASCII ordinal of the character.

The Directory Volume uses the {\em aff2:stored} attribute to provide a
base URL. The URL for each segment is then constructed by appending
the escaped segment URN to the base URL. Note that there is no
restriction on what type of URL this can be, so it may be a location
on a filesystem (e.g. {\em file://some\_directory/}) or a location on a
HTTP server (e.g. {\em http://intranet.server/some\_path}). In this
way its possible to move the entire volume from a filesystem to a web
server transparently.

The Directory Volume stores its own URN in a special segment named
``{\em \_\_URN\_\_}'' at the base of the directory.

\subsection{ZIP64}
For AFF4, we have changed the default volume container file format to
ZIP64\cite{zipspecs}. There are many reasons for this decision:

@@

\item ZIP64 libraries are readily available making proprietary implementations of
interfaces to the AFF4 volume format simple to write. For example, a
simple python program to dump out an Image stream is illustrated in
Listing \ref{python_code}.

@@

\begin{lstlisting}[
	float=tb, label=python_code, caption=Sample Python code to
	dump out an Image Stream. As can be seen the chunk index
	segment is used to slice the data segment into chunks. The
	chunks are decompressed and written to the output file.]
volume=zipfile.ZipFile(INPUT_FILE)
outfd = open(OUTPUT_FILE,"w")
count = 0

while 1:
    idx_segment = volume.read(STREAM+"/%08d.idx" % count)
    bevy = volume.read(STREAM+"/%08d" % count)
    indexes = struct.unpack("<" + "L" * 
	(len(idx_segment)/4), idx_segment)

    for i in range(len(indexes)-1):
	chunk = bevy[indexes[i]:indexes[i+1]]
        outfd.write(zlib.decompress(chunk))

    count += 1
\end{lstlisting}

@@

\begin{figure}[]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
  \caption{The basic structure of a Zip archive}
  \label{zip_structure}
  \end{center}
\end{figure}

It is important to note that AFF4 only requires that the volume be
capable of storing multiple named segments of data. Although our AFF4
implementation uses the ZIP64 file format as an underlying storage
mechanism, our system also supports legacy AFF1 volumes as well as
segments that are stored in other storage systems---for example,
Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.

\subsection{Unused ZIP64 Features}
There are a number of aspects in the Zip file format specifications
which we do not utilize.
\begin{itemize}
\item We ignore Zip64's built-in support for splitting
archives into multiple ZIP files. Instead, our implementation treats
each volume as complete and stand-alone ZIP file. The AFF4
implementation then considers the segments contained within as
belonging to the universal collection. We provide the ability to split
a stream across volumes automatically, as each segment may be stored
in a different volume, and the Universal Resolver is used to find it.

@@

%% Im not sure this is really that relevant here. Byte range updates
%% can be achieved by an append and a map operation, which might be a
%% better option for aimager working in dd_rescue mode.

%\subsubsection{A Forensically Sound Strategy For Updating ZIP64 Files}
%In many forensic applications, individual segments within the volume
%need to be updated. To this end, Garfinkel implemented several
%``update'' functions in AFF1\cite{garfinkel:aff}.

%There were two problems with the original AFF updating approach:
%\begin{itemize}
%\item Updates can cause the volume to
%become fragmented as holes appear within the middle of the volume, and
%updated segments are added at the end.
%\item Updates could be performed silently, raising questions of
%  forensic soundness.
%\end{itemize}

%In AFF4, when a new segment is added to an existing volume, the new
%segment header overwrites the volume central directory, and the
%central directory should be rewritten including the new entry
%immediately after the last file in the archive. If it is not possible
%to overwrite the file, the new segment can be written at the end,
%followed by an updated copy of the central directory (in that case the
%old central directory remains in the file). This is illustrated in
%Figure \ref{zip_structure}. Implementations may delay
%writing the central directory until the file is closed; should the
%system crash before the central directory is written, it can be
%recovered by scanning the entire ZIP file.

%A segment within the archive is considered to be updated when a new
%segment is added to the archive with the same name and a later
%timestamp. Implementations satisfy read requests for duplicate
%segment names with those segments with the later timestamp, or if the
%two timestamps are the same, the segment which appears at a later
%offset in the volume. This effectively allows for a segment to be
%updated without leading to internal volume fragmentation.

%The fact that the older version of all updated segments remains intact
%within each volume can be used for auditing, or to allow a rollback to
%previous versions of the segment. We have created a new tool,
%\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
%creating new volumes and redistributing segments among them.

\section{Streams}
The Stream object provides random access to an abstract representation
of a body of data. Our implementation allows the segments in a stream
to be operated on as if they were a single file by supporting the
traditional POSIX-like functionality of
\texttt{open()}, \texttt{seek()}, \texttt{write()}, and \texttt{read()}.

The following describes a number of types of streams. It is important
to note that clients of our implementation do not care how a
particular stream is implemented. Streams are opened by their URNs,
and the library itself ensure they provide the Stream interface. So
for example, users do not care if a stream is a Map Stream or an Image
Stream - the data provided is the same.

%% Not sure how its relevant - maybe we can move it to a new section
%% about randomly writing stream.

%AFFLIB also supports the \emph{writing} of streams, a feature that is
%used by Garfinkel's \texttt{aimage} disk imaging program. In general
%arbitrary writing to streams can be expensive, especially when writing
%is combined with seek operations. The current \texttt{aimage} program
%will create AFF files with gaps in them during its error recovery
%process (when \texttt{aimage} seeks to the end of a disk and reads it
%backwards in an attempt to get around media failures). We are working
%on a rewrite of \texttt{aimage} which performs imaging and recovery
%on bevy boundaries, rather than using the current arbitrary
%``high-water-mark'' and ``low-water-mark.''

%All streams within an Evidence Set \emph{must} have a properties segment. The
%properties segment describes what the stream contains and determines
%how the stream is processed.

\subsection{The Image Stream}
\label{image_stream}
The AFF4 \emph{Image Stream} stores a single read-only forensic data
set. For example, this stream might contains a hard disk image, a
memory image or a network capture (in PCAP format).

Storage for the data is done by using multiple data segments stored on
various volumes. Data segment URNs are derived by appending an 8
digit, zero padded decimal integer representation of an incrementing
id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032'').

Image streams must specify the
\texttt{page\_size} attribute, as the number of image bytes each chunk
contains. Pages may be compressed at the page level, or may be broken
up into multiple chunks (typically 8KB), and individually
compressed. In this case the beginning of the page segment will
contain an offset table indicating the location of each compression
chunk. This is similar to the technique used by the Expert Witness
file format used by EnCase\cite{encase-3.0} and implemented by the
open source libewf\cite{libewf} package.

@@
